---
title: "econ_assignment3"
author: "Jessica Jagdeo"
date: "May 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, output = FALSE, load necessary packages and read in the data}

library(tidyverse)

library(stargazer)

whale_data <- read_csv("HW3Data.csv")

```

Linear regression
1 = Votes yes
0 = Votes no
    
```{r, do the linear regression}

whales_lm <- lm(vote ~ risk + bid + NEP + income + age, data = whale_data)

summary(whales_lm)


```

###1

Equation:

votingyes = 0.1196977 + 0.0007445(risk) + -0.0010699(bid) + 0.0158639(NEP) + 0.0088282(incomeone_percent) +  0.0027386(incomepoor) + 0.0074891(incomerich) + 0.0467922(incomevery_rich) +  0.0099816(agetofifty) + -0.0201190 (agetoforty) +  -0.0162261(agetosixty) + 0.0204401(agetothirty)

Interpretations:

If everything is held constant, with every percent increase in risk reduction level, a person is about 0.07 times more likely to vote yes. 

###2

Equation:

0.0010699(bid) = 0.1196977 + 0.714votingyes + 0.0007445(risk) + 0.0158639(NEP) + 0.0088282(incomeone_percent) +  0.0027386(incomepoor) + 0.0074891(incomerich) + 0.0467922(incomevery_rich) +  0.0099816(agetofifty) + -0.0201190 (agetoforty) +  -0.0162261(agetosixty) + 0.0204401(agetothirty)


0.0010699(bid) = 0.1196977 + 0.714 + 0.0007445(risk) 

```{r}

#Bid when risk = 0:

bidzero <- (0.1196977 + 0.714 + 0.0007445*0)/0.0010699

#Bid when risk = 4:

bidfour <- (0.1196977 + 0.714 + (0.0007445*4))/0.0010699

bidfour-bidzero


```

###3

```{r}

# Bid/estimated WTP for the 30th respondent using a 60% risk reduction level (voted no):

(0.1196977 + 0.714 + (0.0007445*60) + (0.0158639*25) + (0.0088282*1) - (0.0201190*1))/0.0010699 

# Estimated WTP for the 60th respondent using a 60% risk reduction level (voted yes):

(0.1196977 + 0.714 + (0.0007445*60) + (0.0158639*42) + (0.0088282*1) + (0.0099816*1))/0.0010699 

# Estimated WTP for the 90th respondent using a 60% risk reduction level (voted no):

(0.1196977 + 0.714 + (0.0007445*60) + (0.0158639*46) + (0.0099816*1))/0.0010699

```

